Scalar and Representative Observables for Polarimetric SAR Data

### 2013-11-06 Remote_Sensing_Letter/paper.tex
mostly include Ian's changes

\newsavebox\IBox
\savebox\IBox{\includegraphics[width=2.5in,height=2.5in]{../images/verify_change_ratio_model_on_RADARSAT2_3d.eps}}

@@

\newcommand{\plotWithLegend}[2]{          
          \begin{pspicture}[showgrid=false](\column,\row)% set showgrid=false for the final
	    \rput[bl](0,0){\includegraphics[width=2.5in,height=2.5in]{#1}}% \usebox\IBoxCRThreeR
	    %\rput(5,1.5){\footnotesize{#2}}
	    %\rput{90}(1.5,5){\footnotesize{pdf / histogram}}
	    \psline[linecolor=plot](5,8)(6,8)
	    \psline[linestyle=dashed](5,9)(6,9)%another line style: dotted
	    \rput(8,8){model}
	    \rput(8,9){data}            
          \end{pspicture}
}

@@

\author{
  Thanh-Hai~Le$^{\ast}$$\dag$\thanks{$^\ast$Corresponding
author. Email: le\_thanh\_hai@hotmail.com \vspace{6pt}},
  and Ian~McLoughlin${\ddag}$, 
  and Chan-Hua~Vun${\dag}$\\
\vspace{6pt} $\dag$School of Computer Engineering, Nanyang Technological University, Singapore.\\
$\ddag$ School of Information Science and Technology,
University of Science and Technology of China.\\
\vspace{6pt}\received{November 2013} }

@@

%What is the situation, the background context of your research? the BIG problem!
During the past decades, exponential growth in computing power has allowed the once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since been extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR extends SAR by exploiting the natural polarization property of Electro-Magnetic (EM) waves,
  leading to the availability of multi-channel POLSAR data, compared to traditional one-channel SAR data.

%Research Motivation: Why it is important to solve this problem?
%The POLSAR data, however, is multidimensional and stochastic.
Like SAR data, POLSAR data is stochastic.
Moreover, it is multi-dimensional, making it even harder to interpret.
Under this context, it is therefore important to establish a simple and intuitive understanding of the data.
 Statistical models are undoubtedly crucial in understanding its stochastic nature.
While several %multidimensional 
models have been proposed to describe POLSAR data, they
  unfortunately tend to be complex and not very intuitive due to the multidimensional nature of the data.
Practical POLSAR data processing however, makes heavy use of scalar discrimination measures,
  which should be based on statistically consistent models for scalar and representative observables of the multidimensional data.
It is thus important to establish a scalar and representative observable for this multi-dimensional POLSAR data.

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
There has been a few scalar observables with accompanying statistical models proposal for POLSAR \citep{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all are based on the likelihood ratio statistics.
This statistical test should be based on an exact and consistent statistical distribution
  but so far has only shown to be based on an asymptotic distribution.

%Research Objective: Give your precise problem statement  
This article hence presents a scalar and representative observable, and its associated generic statistical model to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of POLSAR discrimination measures.
%Now State your approach  
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
POLSAR data have been statistically modelled as following the complex Wishart distribution.
Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR data.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

%The specific two and three dimensional models are then validated against both partial and full polarimetric SAR data.
Interestingly the paper also shows that %It is also shown that
  the specific one dimensional versions of the proposed  models 
  match perfectly with the traditional statistical model used for SAR intensity.
This effectively incorporates the common one dimensional SAR theory under the unbrella of the proposed scalar approach for multi dimensional POLSAR.
With this new insight the different discrimination proposals for SAR and POLSAR are reviewed and it is shown that the approach proposed in this paper provides a strong, unifying and consistent foundation relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific one, two and three dimensional versions of the proposed generic models are validated against practical data. 
%Furthermore, the approach proposed in this paper
%  not only provides a consistent foundation for the different discrimination measures proposal in POLSAR
%  but also leads to the establishment of a new discriminating measures: the determinant-ratio and the change-ratio.

@@

%Article Organization: Show that you deliver on the objectives and hence solve the stated problem
The remainder of this article is organized as follows.
After the next section reviews existing discrimination measures as well as scalar observables models for POLSAR, section \ref{sec:theoretical_model}
  derives the generic statistical model for the determinant of the POLSAR covariance matrix
  and proposes several new discrimination measures for the multi-dimensional data.
After section \ref{sec:sar_special_case_of_polsar} illustrates the match between the specific model of one dimensional and the traditional model for SAR intensity,
  section \ref{sec:link_sar_polsar} links the disparate discrimination measure proposals for the traditional SAR and the more recent POLSAR data.
%Section \ref{sec:polsar_models_validation} then validates the specific one, two and three dimensional versions of the proposed model against real-life captured data.
The applicability of these models against practical data is illustrated in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} presents a high-level discussion of the results presented before
  section \ref{sec:conclusion} finally concludes the paper.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in the underlying approach \citep{Conradsen_2003_TGRS_4}.

\subsection{Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
In \citep{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that it is impossible to identify a single best representation.
Although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than providing a unified representation of the POLSAR data.

Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart distribution,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, such derivations are not trivial tasks, and so far, only a handful of such statistical models have been proposed, including the following:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \citep{Joughin_1994_TGRS_562} \citep{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \citep{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \citep{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \citep{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \citep{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \citep{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \citep{Erten_2012_Sensors_2766} have been proposed.

While these models undoubtedly help to further our understandings of the POLSAR data,
  none of the underlying observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and thus 
  2) being representative of the complex POLSAR data.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined respectively as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR, % covariance matrix
  these dissimilarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \citep{Lee_1999_TGRS}.
It is defined \citep{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \citep{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \citep{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \citep{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \citep{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these dissimilarity measures reveals that most of them are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their publications refered the same statistical model developed in \citep{Conradsen_2003_TGRS_4} as the foundation.
In \citep{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above equation, and denoting $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
{\small
\begin{align}
  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{align}
}

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \citep{Conradsen_2003_TGRS_4}, only an asymptotic distribution was used.
This paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

%In this section, the generic scalar statistical model for POLSAR is presented.
In this section, after the basic foundations of POLSAR statistical analysis are introduced,
  the generic scalar statistical model for the multi-dimensional POLSAR will be presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data.

In this paper, the POLSAR scattering vector is denoted as $s$.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

Let $\Sigma=E [ss^{*T}]$ denote the population expected value of the POLSAR covariance matrix,
  where $s^{*T}$ is the complex conjugate transpose of $s$. 
Assuming %all the elements in $s$ are independent and
  $s$ is jointly circular complex Gaussian with the expected covariance matrix $\Sigma$,
  then the probability density function (PDF) of $s$ can be written as:
\begin{equation}
  pdf(s;\Sigma)=\frac{1}{\pi^d|\Sigma|} e^{-s^{*T}\Sigma^{-1}s}
\end{equation}
where $||$ denotes the matrix determinant.

The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
  and $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry.

Complex Wishart distribution statistics are normally used for the scaled covariance matrix
$Z=LC_v$, whose PDF is given as:
\begin{equation}
  pdf(Z;d,\Sigma,L)=\frac{|Z|^{L-d}}{|\Sigma^L|\Gamma_d(L)}e^{-tr(\Sigma^{-1}Z)}
\end{equation}
with $\Gamma_d(L) = \pi^{d(d-1)/2} \prod^{d-1}_{i=0}\Gamma(L-i)$
and $d$ is the dimensional number of the POLSAR covariance matrix.

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \citep{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
This result is used here to develop the generic scalar statistical model for POLSAR. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} %and \ref{eqn:sum_log_chi_squared_rv}
we have: %can deduce the following relationships:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}

>>>>

\newsavebox\IBox
\savebox\IBox{\includegraphics[width=1.5in,height=1.5in]{../images/verify_change_ratio_model_on_RADARSAT2_3d.eps}}

@@

\newcommand{\plotWithLegend}[2]{          
          \begin{pspicture}[showgrid=false](\column,\row)% set showgrid=false for the final
	    \rput[bl](0,0){\includegraphics[width=1.5in,height=1.5in]{#1}}% \usebox\IBoxCRThreeR
	    %\rput(5,1.5){\footnotesize{#2}}
	    %\rput{90}(1.5,5){\footnotesize{pdf / histogram}}
	    \psline[linecolor=plot](5,8)(6,8)
	    \psline[linestyle=dashed](5,9)(6,9)%another line style: dotted
	    \rput(8,8){model}
	    \rput(8,9){data}            
          \end{pspicture}
}

@@

\author{
  Thanh-Hai~Le$^{\ast}$$\dag$\thanks{$^\ast$Corresponding
author. Email: le\_thanh\_hai@hotmail.com \vspace{6pt}},
  and Ian~McLoughlin${\ddag}$, 
  and Chan-Hua~Vun${\dag}$\\
\vspace{6pt} $\dag$School of Computer Engineering, Nanyang Technological University, Singapore.\\
$\ddag$ University of Science and Technology of China.\\
\vspace{6pt}\received{November 2013} }

@@

%What is the situation, the background context of your research? the BIG problem!
Relentless growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
to become a feasible and preferred technique for earth observation.
Basic SAR has also been extended in a few directions, one of which is polarimetric SAR (POLSAR).
This exploits the natural polarization property of Electro-Magnetic (EM) waves,
  encoded in multiple channels, compared to traditional one-channel SAR.

%Research Motivation: Why it is important to solve this problem?
%The POLSAR data, however, is multidimensional and stochastic.
POLSAR data, like SAR, is stochastic but is also multi-dimensional, making it even harder to interpret.
It is therefore important to establish a simple and intuitive understanding of the data.
 Statistical models are undoubtedly crucial in understanding its stochastic nature.
While several %multidimensional 
models have been proposed for POLSAR data, they
   tend to be complex and unintuitive due to the multidimensional nature of the data.
Practical POLSAR data processing however, makes heavy use of scalar discrimination measures,
  which should be based on statistically consistent models %for scalar and representative observables 
  of the multidimensional data.
It is thus important to establish scalar and representative observables for multi-dimensional POLSAR data.

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
A few scalar POLSAR observables with accompanying statistical models have been proposal \citep{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
Alternatively, a few POLSAR discrimination measures have been proposed,
  but all based on likelihood ratio statistics.
This should ideally be based on an exact and consistent distribution
  but so far only asymptotic distributions have been demonstrated.

%Research Objective: Give your precise problem statement  
This article presents a scalar and representative observable, and its associated generic statistical model, to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of discrimination measures.
%Now State your approach  
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
POLSAR data have been statistically modelled as following the complex Wishart distribution.
Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

%The specific two and three dimensional models are then validated against both partial and full polarimetric SAR data.
We will also show that %It is also shown that
  the specific one dimensional (1-D) versions of the proposed  models 
  match %perfectly with 
  the traditional statistical model used for SAR intensity.
This effectively incorporates SAR theory under the umbrella of the proposed scalar approach for multi dimensional POLSAR.
The different discrimination proposals for SAR and POLSAR are reviewed in light of this and the  proposed in this paper is shown to provide a strong, unifying and consistent foundation. % relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific 1, 2 and 3-D versions of the proposed models are validated against practical data. 
%Furthermore, the approach proposed in this paper
%  not only provides a consistent foundation for the different discrimination measures proposal in POLSAR
%  but also leads to the establishment of a new discriminating measures: the determinant-ratio and the change-ratio.

@@

%Article Organization: Show that you deliver on the objectives and hence solve the stated problem

%%***IVM SKIP THIS TO SAVE SPACE - letters usually don't have this
%The remainder of this article is organized as follows.
%After the next section reviews existing discrimination measures as well as scalar observables models for POLSAR, section \ref{sec:theoretical_model}
%  derives the generic statistical model for the determinant of the POLSAR covariance matrix
 % and proposes several new discrimination measures for the multi-dimensional data.
%After section \ref{sec:sar_special_case_of_polsar} illustrates the match between the specific model of one dimensional and the traditional model for SAR intensity,
 % section \ref{sec:link_sar_polsar} links the disparate discrimination measure proposals for the traditional SAR and the more recent POLSAR data.
%The applicability of these models against practical data is illustrated in section \ref{sec:polsar_models_validation}.
%Section \ref{sec:discussion} presents a high-level discussion of the results presented before
 % section \ref{sec:conclusion} finally concludes the paper.

@@

%Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
%  where it is shown that none are able to provide statistically consistent discrimination measures. 
%Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
 % and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
%While an exact statistical distribution is expected to be necessary for a test,
 % only an asymptotic distribution is used in the underlying approach \citep{Conradsen_2003_TGRS_4}.

%**IVM save space!
%\subsection{Scalar Observables for POLSAR Data and their Statistical Models}
%\label{sec:lit_models}

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
\citet{Alberga_2008_IJRS_4129} evaluated the performance of different scalar POLSAR observables for classification.
While many were presented, their corresponding statistical models and classifiers were not available.
Furthermore, the paper concluded that it is impossible to identify a single best representation.
Although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than a unified representation.
%
Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, this is non-trivial -- so far, only a handful of such models have been proposed, including:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \citep{Joughin_1994_TGRS_562} \citep{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \citep{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \citep{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \citep{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \citep{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \citep{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \citep{Erten_2012_Sensors_2766} have been proposed. Although useful, these have not been shown to result in statistically consistent discrimination measures or 
 be representative of the complex POLSAR data.
%While these models undoubtedly help to further our understandings of the POLSAR data,
 % none of the underlying observables have been shown to meet the dual criteria of
 % 1) resulting in statistically consistent discrimination measures and thus 
 % 2) being representative of the complex POLSAR data.

@@

Euclidean or Manhattan distance measures for matrices are not widely used 
for POLSAR due to the multiplicative nature of the noise.
%, defined respectively as:
%\begin{align}
%  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
%  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
%\end{align}
%where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
% $||$ denotes absolute values
%and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
Instead, the Wishart distance is probably most common, as part of the well-known Wishart classifier \citep{Lee_1999_TGRS}, defined \citep{Lee_1994_IJRS_2299} as:
$d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})$
where $tr(C)$ denotes the trace of the POLSAR covariance matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \citep{Anfinsen_2007_ESA_POLINSAR},
\vspace{-2mm}
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
\vspace{-2mm}
the Bartlett distance \citep{Kersten_2005_TGRS_519},
     \vspace{-2mm}
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
     \vspace{-2mm}
the Bhattacharyya distance \citep{Lee_2011_IGARSS_3740},
     \vspace{-2mm}
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
     \vspace{-2mm}
and the Wishart Statistical test distance \citep{Cao_2007_TGRS_3454},
     \vspace{-2mm}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these reveal that most are related:
The Bhattacharyya and Bartlett distances are easily shown to be related.
At the same time, Barlett can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among  measures may be due to the fact that
  all of their publications refer to, and are based on, the same statistical model in \citep{Conradsen_2003_TGRS_4}.
In \citep{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
     \vspace{-2mm}
\begin{equation}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation}

Taking the log-transformation of the above equation, and denoting $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
{\small
\begin{align}
  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{align}
}
\vspace{-2mm}

To detect changes, a test statistic is developed for this discrimination measure.
i.e. a distribution is derived for the dissimilarity measure.
However, \citet{Conradsen_2003_TGRS_4} only use an asymptotic distribution.
By contrast, this paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

In this section, the generic scalar statistical model for POLSAR is presented.
%In this section, after the basic foundations of POLSAR statistical analysis are introduced,
 % the generic scalar statistical model for the multi-dimensional POLSAR will be presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data. 
%
%In this paper, the POLSAR scattering vector is denoted as $s$.
%For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
%  the vector is two-dimensional ($d=2$) and is normally written as: 
%\begin{equation}
%s_{part}=\begin{bmatrix}
%S_h\\ 
%S_v
%\end{bmatrix}
%\end{equation}
%For full and monostatic POLSAR data,
 % the vector is three-dimensional ($d=3$) and is presented as:
%\begin{equation}
%s_{full}=\begin{bmatrix}
%S_{hh}\\
%\sqrt{2}S_{hv}\\
%S_{vv}
%\end{bmatrix}
%\end{equation}
%
First, let $\Sigma=E [ss^{*T}]$ denote the population expected value of the POLSAR covariance matrix,
  where $s^{*T}$ is the complex conjugate transpose of the POLSAR scattering vector, $s$. 
%Assuming 
If this is jointly circular complex Gaussian with expected covariance matrix $\Sigma$,
  then the PDF of $s$ can be written as
$pdf(s;\Sigma)=\{1/{\pi^d|\Sigma|}\} exp\{-s^{*T}\Sigma^{-1}s\}$
where $||$ denotes the matrix determinant.
%
The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
     \vspace{-4mm}
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
 and $s_i$ denotes the partial or full POLSAR scattering vector respectively.
%
Complex Wishart distribution statistics are normally used for the scaled covariance matrix
$Z=LC_v$, whose PDF is given as:
\begin{equation}
  pdf(Z;d,\Sigma,L)=\frac{|Z|^{L-d}}{|\Sigma^L|\Gamma_d(L)}e^{-tr(\Sigma^{-1}Z)}
\end{equation}
with $\Gamma_d(L) = \pi^{d(d-1)/2} \prod^{d-1}_{i=0}\Gamma(L-i)$
and $d$ the dimension number of the POLSAR covariance matrix.
%
The approach taken in this paper differs by applying a homoskedastic log transformation  on a less-than-well-known relationship.
\citet{Goodman_1963_AMS_178} found
that the ratio between observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\vspace{-2mm}
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
We use this to develop a generic scalar statistical model. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} we have:
\vspace{-2mm}
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}
